Word-Referent Identification Under Multimodal Uncertainty

### 2018-04-15 Writeup/ms.tex
modif to ms

How does the listener deal with uncertainty to identify the speaker's
intended word? One rigorous way to approach this question is through
conducting an \emph{ideal observer} analysis. This research strategy
provides a characterization of the task/goal and shows what the optimal
performance should be under this characterization.\footnote{It is, thus,
  a general instance of the rational approach to cognition (Anderson,
  1990). It can also be seens as an instance Marr's computational level
  of analysis.} When there is uncertainty in the input, the ideal
observer performs an optimal probabilistic inference. For example, in
order to recognize an ambiguous linguistic input, the model uses all
available probabilistic knowledge in order to maximize the accuracy of
this recognition. When the task is well specified, the ideal observer
model can be seen as a theoretical upper limit on performance. Thus, it
is supposed to be used, not so much as a realistic model of human
performance, as much as a baseline against which human performance can
be compared (Geisler, 2003; Rahnev \& Denison, 2018). When there is a
deviation from the ideal, it can reveal extra constraints on human
cognition, such as limitations on the working memory or attentional
resources.

@@

On the face of it, the question of combining information from the sound
and the visual referent might seem similar to that of audio-visual
speech integration. Nevertheless, there are at least two fundamental
differences between these two cases, and both can influence the way the
auditory and visual cues are integrated:

\noindent First, in the case of audio-visual speech, both modalities
offer information about the same underlying speech category. They may
differ only in terms of their informational reliability. In a
referential context, however, the auditory and visual modalities are
additionally different in terms of the roles they play in the
referential process: the auditory input represents the \emph{symbol}
whereas the visual input represents the \emph{meaning}. It has been
suggested that because of its referential property, speech is a
privileged signal for humans, starting in infancy (see Vouloumanos \&
Waxman, 2014 for a review).\footnote{There is a debate as to whether
  speech is privileged for infant and adults for the same reasons.
  Whereas some researchers suggest that speech is privileged for both
  infants and adults because of its ability to refer (e.g., Waxman \&
  Markow, 1995), others suggest that speech might \emph{not} have a
  referential status from the start. Rather, speech might be prefered by
  infants only because of a low level auditory ``overshadowing'' (e.g.,
  Sloutsky \& Napolitano, 2003).} Thus, in a referential context, it is
possible that listeners do not treat the auditory and visual modalities
as equivalent sources of information. Instead, there could be a
sub-optimal bias for the auditory modality beyond what is expected from
informational reliability alone.

@@

We construct an ideal observer model that combines probabilistic
information from the auditory and visual modalities. In contrast to the
model used in most research on multisensory integration (e.g., Ernst \&
Banks, 2002)---which typically studies continuous stimuli (e.g., size,
location)---this probabilistic information cannot be characterized with
\emph{sensory noise/variability}, only. Indeed, our task involves
responses over categorical variables (phonemes and concepts), and
therefore, the optimal model in our case should take into account, not
only the noise variability around an individual perceptual estimate, but
also its \emph{within category variability}, i.e., the uncertainty
related to whether this perceptual estimate belongs to a given category
(see also Bankieris, Bejjanki, \& Aslin, 2017; Bejjanki et al., 2011).
In what follows, we describe a model that accounts for both type of
variability. First, we describe the model in the simplified case of
categorical variability only. Second, we augment this simplified model
to account for sensory noise.

@@

The optimal model (as well as the auditory and visual baselines) are
\emph{normative} models. Their predictions are made about human data in
the bimodal condition, but their crucial parameters (i.e., variances
associated with the visual and auditory modalities) are derived from
data in the unimodal conditions. In addition to these normative models,
we consider a \emph{descriptive} model. The parameters of this model are
fit to actual responses in the bimodal condition. If the referential
task induces sub-optimality (due, for instance, to the arbitrary nature
of the sound-object association), then the descriptive model should
explain more variance than the optimal model does.

@@

There are three possible ways human responses can deviate from
optimality. These scenarios are illustrated in Figure~\ref{fig:subOptim}
and are as follows:

@@

As a second analysis, we asked whether the observed variance in the
individual distribution was due to mere sampling errors or whether it
corresponded to a real between-subject variability. We simulated
individual responses from the posterior distribution whose parameters
were fit to the population as a whole (i.e., the descriptive posterior).
The resulting distribution is shown in Figure~\ref{fig:individual}. For
ease of comparison, the simulated distribution was superimposed to the
real distribution. We found that the real distribution was broader
(\(sd=\) 2.24) than the simulated distribution (\(sd=\) 1.17),
indicating that there was a real between-subject variation beyond
sampling errors. This means that the participants varied in terms of how
they weighted modalities: Compared to the predictions of the global
descriptive model, some participants relied more on the auditory
modality, whereas others relied more on the visual modality.

@@

The percentage of participants who relied exclusively on either the
visual modality or the auditory modality was 38.64\%, which is closer to
the percentage of Experiment 2, except that now almost all of them
relied on the auditory modality (94.12\%). For ease of comparison, Table
\ref{tab:exclusive} provides a summary of the numbers across the three
Experiments.

@@

When identifying a spoken word under uncertainty, one often needs to
make the most of the available cues. Some previous work studied optimal
behavior under uncertainty from the auditory input only (e.g., Clayards
et al., 2008; Feldman et al., 2009), and others studied optimality under
multimodal uncertainty in auditory speech and visual facial features
(e.g. Bejjanki et al., 2011). The current work explored, for the first
time, the case of word identification under uncertainty in speech (word
form) and the visual word \emph{referent}. More specifically, we
conducted an ideal observer analysis of the task whereby a model
provided predictions about how information from each modality should be
combined in an optimal fashion. The predictions of the model were tested
in a series of three experiments where instances of both the form and
the meaning were ambiguous with respect to their category membership
only (Experiment 1), when instances of the form were perturbed with
additional background noise (Experiment 2), and when instances of the
referent were perturbed with additional visual noise (Experiment 3).

@@

The second case of sub-optimality is related to how participants weighed
the cues from the visual and the auditory modalities in a noisy context.
In contrast to Experiment 1 where the combination was indistinguishable
from the optimal prediction, results of Experiment 2 and 3 which both
involved background noise in one modality, showed that participants had
a systematic preference for the other (non-noisy) modality. From
previous empirical studies, we know that when the speech signal is
degraded, people tend to compensate by relying more on other sources of
information such as the accompanying visual cues (i.e., lip movements)
or the semantic/syntactic context (see Mattys, Davis, Bradlow, \& Scott,
2012 for a review). However, and generally speaking, these studies do
not differentiate between an optimal compensatory strategy (i.e.,
relying more on the alternative source while using all information still
available in the distorted signal), and a sub-optimal strategy (i.e.,
relying more on the alternative source while ignoring at least some of
the information still available in the distorted signal). The formal
approach followed in this paper allowed us to tease apart these two
possibilities, and the analysis supports the sub-optimal compensatory
strategy: The preference for the non-noisy modality is above and beyond
what can be explained by the relative reliability alone, meaning that
the participants tend to ignore at least part of the information still
available in the noisy modality.

@@

Though we only tested adults in this paper, the problem of word
recognition under uncertainty, as well as the need to make the most of
ambiguous cues, is a particularly pressing issue for children. In fact,
whereas adults are mostly faced with uncertainty in the \emph{input},
children have to deal with the additional uncertainty that results from
their early unrefined \emph{representations} of both phonological and
semantic categories. For example, upon hearing a noisy instance of
\enquote{bee}, adults may have to decide whether the speaker intended to
say \enquote{pea} or \enquote{bee}, but children can additionally be
uncertain whether \enquote{bee} is a different word from \enquote{pee}
(as opposed to, say, a valid within category variation), especially if
these similar sounding words are newly learned (Creel, 2012; Merriman \&
Schuster, 1991; Stager \& Werker, 1997; Swingley, 2016; White \& Morgan,
2008). Though new word form representation can be shown to be
differentiated under some circumstances (e.g., Yoshida, Fennell,
Swingley, \& Werker, 2009), this differentiation is still not mature
enough and is probably noisier than the adult-like representation and/or
encoded with lower confidence (see Swingley, 2007).

@@

Thus, uncertainty in the representation associated with one modality
(e.g., a bee and a fly) can be mitigated through the possibly more
differentiated representations associated with the other modality (e.g.,
the sound \enquote{bee} is acoustically different from the sound
\enquote{fly}). That being said, a multi-modal cue combination strategy
might help children not only recognize an individual word instance, but
also refine the underlying phonological and semantic representations in
the process. Previous research in early word learning has---whether
implicitly or explicitly---largely treated the process of learning form
and of learning meaning as independent. However, the developmental data
reviewed above shows that children do not wait to have completed the
acquisition of form to start learning meanings, and that both form and
meaning representations develop, rather, in a parallel fashion. A few
studies pointed to the possibility of an interaction between sound and
meaning in early acquisition. For instance, Waxman and Markow (1995)
showed that labeling various objects with the same name helps infants
form the broad semantic category (but see Sloutsky \& Napolitano, 2003).
Vice versa, Yeung and Werker (2009) showed that pairing similar sounds
with different objects help infants pay attention to subtle phonological
contrasts. The present study proposes a first step towards a formal
framework where accounts of sound-meaning synergy in development can be
unified and further explored. For example, one could imagine that,
initially, visual and auditory categories have relatively large noise
variances, and that development consists in reducing the values of the
variances through a mutually constraining process as further multimodal
data accumulates.

One limitation of the work is that we used simplified stimuli. For the
auditory modality, we used speech categories that varied along a single
acoustic dimension. While this dimension might be sufficient to
recognize words in our specific case, in general the speech signal may
be more complex, varying along several acoustic/phonetic dimensions.
Additionally, these dimensions may be highly variable due to various
kinds of speaker and context differences. The same thing can be said
about the referential stimuli. Here we used a continuum along a single
morph dimension in order to construct a multimodal input where the
auditory and visual components have symmetrical properties. Though such
morph is not the exact visual variability that people would encounter in
they daily lives, it allowed us to precisely test the role of auditory
and visual information in the cue combination process. Parameterizing
semantic dimensions is a notoriously difficult problem, but morphs have
been used in previous research as a reasonable proxy (Freedman et al.,
2001; Havy \& Waxman, 2016; Sloutsky \& Fisher, 2004). It is an open
question as to whether people use the same strategy in controlled
laboratory conditions, as in more naturalistic settings where they have
to deal with various levels of variability. However an answer to this
question is likely to involve a multifaceted approach,
involving---besides laboratory experiments---analyses of corpora with a
more realistic multimodal input (e.g., Fourtassi, Schatz, Varadarajan,
\& Dupoux, 2014; Harwath, Torralba, \& Glass, 2016; B. C. Roy, Frank,
DeCamp, \& Roy, 2015).

In conclusion, this work studied the mechanism of word identification
under uncertainty in both the word form and the word referent. To our
knowledge, this is the first study that performs an ideal observer
analysis of this task. We found people to be near optimal in their cue
combination: They weighted each modality according to its relative
reliability. However, they also showed patterns of sub-optimality
especially when the stimuli were perturbed with additional background
noise. Though the present study did not directly address the issue of
early word learning, it provides a framework where developmental
questions can also be investigated. For instance, future work should
explore whether children, like adults, use probabilistic cues from both
the auditory and the visual input to recognize ambiguous words, the
extent to which they combine these cues in an optimal fashion, and
whether these combination help them with refining their early
phonological and semantic representations.

>>>>

How does the listener deal with uncertainty to identify the speaker's
intended word? One rigorous way to approach this question is through
conducting an \emph{ideal observer} analysis. This research strategy
provides a characterization of the task/goal and shows what the optimal
performance should be under this characterization.\footnote{It is, thus,
  a general instance of the rational approach to cognition (Anderson,
  1990). It can also be seens as an instance of Marr's computational
  level of analysis.} When there is uncertainty in the input, the ideal
observer performs an optimal probabilistic inference. For example, in
order to recognize an ambiguous linguistic input, the model uses all
available probabilistic knowledge in order to maximize the accuracy of
this recognition. When the task is well specified, the ideal observer
model can be seen as a theoretical upper limit on performance. It is
supposed to be used, not so much as a realistic model of human
performance, as much as a baseline against which human performance can
be compared (Geisler, 2003; Rahnev \& Denison, 2018). When there is a
deviation from the ideal, it can reveal extra constraints on human
cognition, such as limitations on the working memory or attentional
resources.

@@

On the face of it, the question of combining information from the sound
and the visual referent might seem similar to that of audio-visual
speech integration. Nevertheless, there are at least two fundamental
differences between these two cases, and both can influence the way the
auditory and visual cues are combined:

\noindent First, in the case of audio-visual speech, both modalities
offer information about the same underlying speech category. They may
differ only in terms of their informational reliability. In a
referential context, however, the auditory and visual modalities are
additionally different in terms of the roles they play in the
referential process: the auditory input represents the \emph{symbol}
whereas the visual input represents the \emph{meaning}. It has been
suggested that because of its referential property, speech is a
privileged signal for humans, starting in infancy (see Vouloumanos \&
Waxman, 2014 for a review).\footnote{There is a debate as to whether
  speech is privileged for children and adults for the same reasons.
  Whereas some researchers suggest that speech is privileged for both
  children and adults because of its ability to refer (e.g., Waxman \&
  Markow, 1995), others suggest that speech might \emph{not} have a
  referential status from the start. Rather, speech might be prefered by
  children only because of a low level auditory ``overshadowing'' (e.g.,
  Sloutsky \& Napolitano, 2003).} Thus, in a referential context, it is
possible that listeners do not treat the auditory and visual modalities
as equivalent sources of information. Instead, there could be a
sub-optimal bias for the auditory modality beyond what is expected from
informational reliability alone.

@@

We construct an ideal observer model that combines probabilistic
information from the auditory and visual modalities. In contrast to the
model used in most research on multisensory integration (e.g., Ernst \&
Banks, 2002)---which typically studies continuous stimuli (e.g., size,
location)---this probabilistic information in our case cannot be
characterized with \emph{sensory noise/variability}, only. Indeed, our
task involves responses over categorical variables (phonemes and
concepts), and therefore, the optimal model in our case should take into
account, not only the noise variability around an individual perceptual
estimate, but also its \emph{categorical variability}, i.e., the
uncertainty related to whether this perceptual estimate belongs to a
given category (see also Bankieris, Bejjanki, \& Aslin, 2017; Bejjanki
et al., 2011). In what follows, we describe a model that accounts for
both type of variability. First, we describe the model in the simplified
case of categorical variability only. Second, we augment this simplified
model to account for sensory noise.

@@

The optimal model (as well as the auditory and visual baselines) are
\emph{normative} models. Their predictions are made about human data in
the bimodal condition, but their crucial parameters (i.e., variances
associated with the visual and auditory modalities) are derived from
data in the unimodal conditions. In addition to these normative models,
we consider a \emph{descriptive} model. It is formally identical to the
normative optimal model (Equation 1), except that the parameters are fit
to actual responses in the bimodal condition. If the referential task
induces sub-optimality (due, for instance, to the arbitrary nature of
the sound-object association), then the descriptive model should explain
more variance than the optimal model does.

@@

There are three possible ways human responses can deviate from
optimality. These scenarios are illustrated in
Figure~\ref{fig:subOptim}, and are as follows:

@@

As a second analysis, we asked whether the observed variance in the
individual distribution was due to mere sampling errors or whether it
corresponded to a real between-subject variability. We simulated
individual responses from the posterior distribution whose parameters
were fit to the population as a whole (i.e., the descriptive posterior).
The resulting distribution is shown in Figure~\ref{fig:individual}. For
ease of comparison, the simulated distribution was superimposed to the
real distribution. We found that the real distribution was broader
(\(sd=\) 2.24) than the simulated distribution (\(sd=\) 1.19),
indicating that there was a real between-subject variation beyond
sampling errors. This means that the participants varied in terms of how
they weighted modalities: Compared to the predictions of the global
descriptive model, some participants relied more on the auditory
modality, whereas others relied more on the visual modality.

@@

The percentage of participants who relied exclusively on either the
visual modality or the auditory modality was 38.64\%, which is closer to
the percentage of Experiment 2, except that now almost all of them
relied on the auditory modality (94.12\%). For ease of comparison, Table
\ref{tab:exclusive} provides a summary of the numbers across the three
experiments.

@@

When identifying a spoken word under uncertainty, one often needs to
make the most of the available cues. Some previous work studied optimal
behavior under uncertainty from the auditory input only (e.g., Clayards
et al., 2008; Feldman et al., 2009), and others studied optimality under
multimodal uncertainty in auditory speech and visual facial features
(e.g. Bejjanki et al., 2011). The current work explored, for the first
time, the case of word identification under uncertainty in speech (word
form) and the visual \emph{referent}. More specifically, we conducted an
ideal observer analysis of the task whereby a model provided predictions
about how information from each modality should be combined in an
optimal fashion. The predictions of the model were tested in a series of
three experiments where instances of both the form and the meaning were
ambiguous with respect to their category membership only (Experiment 1),
when instances of the form were perturbed with additional background
noise (Experiment 2), and when instances of the referent were perturbed
with additional visual noise (Experiment 3).

@@

The second case of sub-optimality is related to how participants
weighted the cues from the visual and the auditory modalities in a noisy
context. In contrast to Experiment 1 where the combination was
indistinguishable from the optimal prediction, results of Experiment 2
and 3 which both involved background noise in one modality, showed that
participants had a systematic preference for the other (non-noisy)
modality. From previous empirical studies, we know that when the speech
signal is degraded, people tend to compensate by relying more on other
sources of information such as the accompanying visual cues (i.e., lip
movements) or the semantic/syntactic context (see Mattys, Davis,
Bradlow, \& Scott, 2012 for a review). However, and generally speaking,
these studies do not differentiate between an optimal compensatory
strategy (i.e., relying more on the alternative source while using all
information still available in the distorted signal), and a sub-optimal
strategy (i.e., relying more on the alternative source while ignoring at
least some of the information still available in the distorted signal).
The formal approach followed in this paper allowed us to tease apart
these two possibilities, and the analysis supports the sub-optimal
compensatory strategy: The preference for the non-noisy modality is
above and beyond what can be explained by the relative reliability
alone, meaning that the participants tend to ignore at least part of the
information still available in the noisy modality.

@@

Though we only tested adults in this paper, the problem of word
recognition under uncertainty, as well as the need to make the most of
ambiguous cues, is a particularly pressing issue for children. In fact,
whereas adults are mostly faced with uncertainty in the \emph{input},
children have to deal with the additional uncertainty that results from
their early unrefined \emph{representations} of both phonological and
semantic categories. For example, upon hearing a noisy instance of
\enquote{bee}, adults may have to decide whether the speaker intended to
say \enquote{pea} or \enquote{bee}, but children can additionally be
uncertain whether \enquote{bee} is a different word from \enquote{pee}
(as opposed to, say, a valid within category variation), especially if
these similar sounding words are newly learned (Creel, 2012; Merriman \&
Schuster, 1991; Stager \& Werker, 1997; Swingley, 2016; White \& Morgan,
2008). Though similar word form representation can be shown to be
differentiated under some circumstances (e.g., Yoshida, Fennell,
Swingley, \& Werker, 2009), this differentiation is still not mature
enough and is probably noisier than the adult-like representation and/or
encoded with lower confidence (see Swingley, 2007).

@@

Thus, uncertainty in the representation associated with one modality
(e.g., a bee and a fly) can be mitigated through the possibly more
differentiated representations associated with the other modality (e.g.,
the sound \enquote{bee} is acoustically different from the sound
\enquote{fly}). That being said, a multi-modal cue combination strategy
might help children not only recognize an individual word instance, but
also refine the underlying phonological and semantic representations in
the process. Previous research in early word learning has---whether
implicitly or explicitly---largely treated the process of learning form
and of learning meaning as independent. However, the developmental data
reviewed above shows that children do not wait to have completed the
acquisition of form to start learning meanings, and that both form and
meaning representations develop, rather, in a parallel fashion. A few
studies pointed to the possibility of an interaction between sound and
meaning in early acquisition. For instance, Waxman and Markow (1995)
showed that labeling various objects with the same name helps infants
form the broad semantic category (but see Sloutsky \& Napolitano, 2003).
Vice versa, Yeung and Werker (2009) showed that pairing similar sounds
with different objects help infants pay attention to subtle phonological
contrasts. The present study proposes a first step towards a formal
framework where isolated accounts of sound-meaning interaction in
development can be unified and further explored.

One limitation of this work is that we used simplified stimuli. For the
auditory modality, we used speech categories that varied along a single
acoustic dimension. While this dimension might be sufficient to
recognize words in our specific case, in general the speech signal may
be more complex, varying along several acoustic/phonetic dimensions.
Additionally, these dimensions may be highly variable due to various
kinds of speaker and context differences. The same thing can be said
about the referential stimuli. Here we used a continuum along a single
morph dimension in order to construct a multimodal input where the
auditory and visual components have symmetrical properties. Though such
morph is not the exact visual variability that people would encounter in
their daily lives, it allowed us to precisely test the role of auditory
and visual information in the cue combination process. Parameterizing
semantic dimensions is a notoriously difficult problem, but morphs have
been used in previous research as a reasonable proxy (Freedman et al.,
2001; Havy \& Waxman, 2016; Sloutsky \& Fisher, 2004). It is an open
question as to whether people use the same strategy in controlled
laboratory conditions, as in more naturalistic settings where they have
to deal with various levels of variability. However an answer to this
question is likely to involve a multifaceted research approach,
involving---besides laboratory experiments---analyses of corpora with a
more realistic multimodal input (e.g., Fourtassi, Schatz, Varadarajan,
\& Dupoux, 2014; Harwath, Torralba, \& Glass, 2016; B. C. Roy, Frank,
DeCamp, \& Roy, 2015).

\section{Conclusion}\label{conclusion}

This work studied the mechanism of word identification under uncertainty
in both the word form and the word referent. To our knowledge, this is
the first study that performs an ideal observer analysis of this task.
We found people to be near optimal in their cue combination: They
weighted each modality according to its relative reliability. However,
they also showed patterns of sub-optimality especially when the stimuli
were perturbed with additional background noise. Though the present
study did not directly address the issue of early word learning, it
provides a framework where developmental questions can also be
investigated. For instance, future work should explore whether children,
like adults, use probabilistic cues from both the auditory and the
visual input to recognize ambiguous words, the extent to which they
combine these cues in an optimal fashion, and whether these combination
help them with refining their early phonological and semantic
representations.
